Constraining gravitational wave amplitude birefringence with GWTC-3

### 2022-08-09 src/tex/ms.tex
Update ms.tex

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space from beyond-GR theory
            \item birefringence: enhance one and suppress another
            \item GR predicts there is no birefringence
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item Assume GW are generated as GR described
        \item birefringence is considered during the propagation
        \item 
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item 
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item Better data from detector
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

\section{Everything}
Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that gravitational waves (GWs) amplitude birefringence is a property of space;
in contrast, GR predicts there is no birefringence.

GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
In GR, the amplitude ratio of the left and right polarisations only depends on the inclination $\iota$ for a nonprecessing binary,
the angle between our line of sight and the orbital angular momentum of the binary.
\begin{equation}
    \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}where $h_L$ and $h_R$ are the Fourier amplitude of left and right polarizations of the GWs respectively.

For other studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
but also the distance to the merger and the strength of the birefringence.
\begin{equation}
    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
\end{equation}where $d_C$ is the comoving distance to the binary, $\widetilde{\kappa}$ is the opacity parameter that represent
the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.

@@

Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog, GWTC-3, was performed to constrain the plausible range of $\kappa$.
Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from
interferometer and priors of the parameters provided. \citep{Ashton_2019}

@@

\begin{figure*}[h]
    \script{GW150914_corner.py}
    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the parameter estimations (PE) done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GWs amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure*}

>>>>

\begin{enumerate}
    \item Introduction
    \begin{enumerate}
        \item Motivation
        \begin{enumerate}
            \item GR broke down at some scale, find a new theory
            \item CS: beyond-GR theory
            \item test CS property (GW amplitude birefringence)
        \end{enumerate}
        \item Amplitude Birefringence
        \begin{enumerate}
            \item GW amplitude birefringence is a property of space from beyond-GR theory
            \item birefringence: enhance one and suppress another
            \item GR predicts there is no birefringence
            \item GR: amplitude ratio depends on $\iota$ only
            \item nonGR: also depends on $\kappa$, $d_C$ and $f$
        \end{enumerate}
    \end{enumerate}
    \item Method
    \begin{enumerate}
        \item Assume GW are generated as GR described
        \item birefringence is considered during the propagation
        \item 
    \end{enumerate}
    \item Results
    \begin{enumerate}
        \item GW150914
        \begin{enumerate}
            \item 
        \end{enumerate}
        \item GWTC-3
        \begin{enumerate}
            \item 
        \end{enumerate}
    \end{enumerate}
    \item Discussion
    \begin{enumerate}
        \item Limitation
        \begin{enumerate}
            \item orientation of detector (polarisations)
            \item sensitivity of detector
        \end{enumerate}
        \item Future Study
        \begin{enumerate}
            \item More events from detector
            \item Higher order terms
            \item Longer sampling
        \end{enumerate}
    \end{enumerate}
\end{enumerate}

\section{Everything}
Since Einstein proposed his theory of general relativity (GR), it was tested in a wide range of length scales.
After a century, we know that GR do not agree with quantum theories at some length scale.
To unify both theories, we need to study the possibility of different beyond-GR theories.
Some beyond-GR theories such as Chern-Simons theory (CS) suggested that there is gravitational wave (GW) amplitude birefringence,
while GR predicts there is no birefringence.

GW amplitude birefringence is a property of space which consists of the enhancement of one GW polarisation over the other during the propagation of the wave.
To illustrate the property, \textbf{[use snapshot from the animation to explain both frequency-dependent and frequency-independent birefringence]}

Before we discuss more about GW amplitude birefringence, we need to understand the case of GR.
GW consist of two linear polarizations (i.e. $+$ and $\times$) similar to electromagnetic waves,
which could be transformed into two circular polarizations (i.e. left and right) by $h_{\mathrm{R}, \mathrm{L}} = h_+ \pm i h_\times$.
In GR, the amplitude ratio of the left and right polarisations only depends on the inclination $\iota$ for a nonprecessing binary,
the angle between our line of sight and the orbital angular momentum of the binary.
\begin{equation}
    \left(\frac{h_\mathrm{L}}{h_\mathrm{R}}\right)_\mathrm{GR}=\left(\frac{1-\cos\iota}{1+\cos\iota}\right)^2\,,
\end{equation}where $h_L$ and $h_R$ are the Fourier amplitude of left and right polarizations of the GWs respectively.

For other studies on the birefringence property such as \citet{Maria_2021}, the amplitude ratio depends on not only $\iota$,
but also the distance to the binary and the strength of the birefringence.
\begin{equation}
    \left(\frac{h_\mathrm{L_{obs}}}{h_\mathrm{R_{obs}}}\right)_\mathrm{Biref}^\mathrm{old}=\frac{e^{d_C\widetilde{\kappa}}\left(1-\cos\iota\right)^2}{e^{-d_C\widetilde{\kappa}}\left(1+\cos\iota\right)^2}\,,
\end{equation}where $d_C$ is the comoving distance to the binary, $\widetilde{\kappa}$ is the opacity parameter that represent
the strength of the birefringence with units of $L^{-1}$, and $h_\mathrm{L_{obs}}$ and $h_\mathrm{R_{obs}}$ are
the observed Fourier amplitude of left and right polarizations of the GW respectively. Note that GR is recovered if $\kappa$ is 0.
This is a zeroth-order approximation of the birefringence model in CS, which assume there is no frequency dependence.
This would create a degeneracy between kappa and iota, as they can affect the amplitude ratio in the same way.

@@

Estimations on $\kappa$ using Bilby with data from the third LIGO-Virgo catalog \textbf{[cite LIGO]}, GWTC-3,
was performed to constrain the plausible range of $\kappa$.
Bilby is a bayesian toolkit for GW analysis, which could calculate posterior of GW waveform parameters based on signals from
interferometer and priors of the parameters provided. \citep{Ashton_2019}

@@

\begin{figure*}[h]
    \script{GW150914_corner.py}
    \includegraphics[width=\textwidth]{figures/GW150914_corner.png}
    \caption{The posterior of $\kappa$, luminosity distance $d_L$ and $\cos{\iota}$ for GW150914.
    The three sets of plots are the parameter estimations (PEs) done by LIGO with GR \citep{LIGO_2021},
    the PE done by us with the frequency independent birefringence and the frequency dependent birefringence respectively.
    Note that there is no posterior of $\kappa$ for the PE from LIGO, as the LIGO PE is based on GR,
    which does not suggest GW amplitude birefringence.}
    \label{fig:GW150914_corner}
\end{figure*}
